\texttt{yacoub}: a Python package for Simulating Generalized Fading Channels

### 2018-03-17 paper/yacoub.tex
update paper

\begin{abstract}
    We present a well tested Python-based library for simulating and computing
    generalized fading channels, named \texttt{yacoub}. We describe the
    applicability of \texttt{yacoub} using examples in recent communications
    systems challenges, namely: spectrum sensing, bit error rate computation,
    and parameter estimation in generalized fading channels. The development of
    \texttt{yacoub} open source and its code is avaliable at
    \texttt{http://github.com/mirca/yacoub}.
\end{abstract}

@@

\subsection{Note on notation}
Scalars and random variables are denoted as \textit{italic} small-case letters \textit{e.g.} $x$;
vectors and random vectors are denoted as \textit{italic}, boldface, small-case letters \textit{e.g.} $\bm{x}$.
The $n-$th component of a vector $\bm{x}$ is denoted as $x_n$.
A complex vector of length $n$ is defined as $\bm{x} \in \mathbb{C}^{n\times 1}$. All
vectors are column vectors. Matrices are denoted as \textit{italic}, boldface, capital
letters as in $\bm{X}$; the identity matrix of order $n$ is denoted as $\bm{I}_n$.
We define a discrete-time circularly symmetric Gaussian process $\bm{z}$ as any collection
of random varibles $\bm{z}~=~\bm{x}~+~j\bm{y}$, $j \triangleq \sqrt{-1}$, such that
$\bm{x}$ and $\bm{y}$ are iid jointly Gaussian with zero mean vector and covariance
matrix given by $\mathbb{E}\left[\bm{z}\bm{z}^{\dagger}\right]$, in which $\bm{z}^\dagger$
means the conjugate transpose of $\bm{z}$. The expectation value wrt a random variable $x$
is denoted as $\mathbb{E}_x$. The probability of an event $A$ is denoted as $\mathbb{P}(A)$.
The indicator function is denoted as $\mathbb{I}(\cdot)$ and it evaluates to $1$ if its
argument is true and $0$ otherwise.

@@

Note that,
\begin{align}
    -\log\Gamma(m) \geq &-\log\Gamma(m_t) + (m - m_t) \nonumber \\
                        &+ \left(m_t - \dfrac{1}{2}\right)\log (m_t-1)\nonumber\\
                        &-\left(m - \dfrac{1}{2}\right)\log (m-1), m \geq m_t > 1
\end{align}
and
\begin{align}
    &m \log m \geq m\log m_t + m - m_t,~m > 0\\
    & - m\log (m-1) \geq - \log(m-1) - (m-1)\left(\log (m_t-1) + \dfrac{m-1}{m_t-1} - 1\right),~m > 1
\end{align}
(in all inequalities, equality is achieved iff $m = m_t$).

Therefore,
\begin{figure*}[!htb]
\begin{align}
    - \log p(\bm{h}) \leq -n&\left(m\log m_t + 2(m - m_t) - m\log\Omega - \log\Gamma(m_t) +
        \left(m_t - \dfrac{1}{2}\right)\log (m_t-1) - \dfrac{1}{2}\log(m-1)\right. \nonumber\\
        &\left.-(m-1)\left(\log (m_t-1) + \dfrac{m-1}{m_t-1} - 1\right)
        \right) +m\left(\dfrac{\sum_{i=1}^{n}h_i^2}{\Omega} - 2\sum_{i=1}^{n}\log h_i\right), m \geq m_t > 1.
\end{align}
\end{figure*}

\begin{figure*}[!htb]
    \begin{align}
    \dfrac{\partial g(m | m_t)}{\partial m} = -n\left(\log m_t + 2 - \log\Omega - \dfrac{1}{2(m - 1)}
    - \left(\log (m_t - 1) - 1\right) + \dfrac{-2m + 1}{m_t - 1} \right)
    + \dfrac{\sum_{i=1}^{n}h_i^2}{\Omega} - 2\sum_{i=1}^{n}\log h_i = 0
    \end{align}
\end{figure*}

\begin{figure*}[!htb]
    \begin{align}
        2(m - 1) \left(-n\left(\log m_t + 2 - \log\Omega - (\log(m_t - 1) - 1)\right) + \dfrac{\sum_{i=1}^{n}h_i^2}{\Omega} - 2\sum_{i=1}^{n}\log h_i \right) + n - \dfrac{2n}{m_t-1}(-2m^2 + 3m - 1) = 0\\
        2(m - 1) a + n - \dfrac{2n}{m_t-1}(-2m^2 + 3m - 1) = 0\\
        m^2\left(\dfrac{4n}{m_t - 1}\right) + m\left(\dfrac{6n}{m_t - 1} + 2a\right) -2a + n\left(\dfrac{m_t +1}{m_t - 1}\right) = 0\\
    \end{align}
\end{figure*}

\subsection{Parameter Estimation in $\alpha$-$\mu$ fading}

The $\alpha$-$\mu$ fading probability density is given as
\begin{align}
p(h) = \dfrac{\alpha \mu^\mu h^{\alpha\mu -1}}{\Gamma(\mu)}\exp\left(-\mu h^\alpha\right),
\end{align}
$h \in \mathbb{R}_{+}$, $\alpha \in \mathbb{R}_{+}$, and $\mu \in \mathbb{R}_{+}$.

Given a vector $\bm{h} \in \mathbb{R}^{n \times 1}$ of iid samples from the
$\alpha-\mu$ distribution, we would like to estimate the value of...
The pdf of $\bm{h}$ is given as
\begin{align}
    p(\bm{h}) = \prod_{i=1}^{n}p(h_i) = \prod_{i=1}^{n}\dfrac{\alpha \mu^\mu h_i^{\alpha\mu -1}}
    {\Gamma(\mu)}\exp\left(-\mu h_i^\alpha\right)\\
    = \left(\dfrac{\alpha \mu^\mu }{\Gamma(\mu)}\right)^{n}
    \exp\left(-\mu\sum_{i=1}^{n}h_i^{\alpha}\right)\prod_{i=1}^{n}h_i^{\alpha\mu -1}
\end{align}
and the loglikelihood is given as
\begin{align}
    \log p(\bm{h}) \cong
    &~n\left(\log\alpha + \mu \log \mu - \log \Gamma(\mu)\right) \nonumber\\
    & - \mu\sum_{i=1}^{n}h_i^{\alpha} + \alpha\mu\sum_{i=1}^{n}\log h_i
\end{align}

Therefore, we want to solve
\begin{align}
    & \text{minimize} -\log p(\bm{h})\\
    & \text{s.t.}~\alpha > 0, \bm{h} > 0 \nonumber
\end{align}

@@

>>>>

\begin{abstract}
    We present a well tested Python-based library for simulating and computing
    generalized fading channels, named \texttt{yacoub}. We describe the
    applicability of \texttt{yacoub} using examples in recent communications
    systems challenges, namely: spectrum sensing, bit error rate computation,
    and parameter estimation in generalized fading channels. For the latter,
    we develop an iterative algorithm based on the Majorization-Minimization
    framework. The development of \texttt{yacoub} is open source and its code
    is avaliable at \texttt{http://github.com/mirca/yacoub}.
\end{abstract}

@@

\subsection{Note on notation}
Scalars and random variables are denoted as \textit{italic} small-case letters \textit{e.g.} $x$;
vectors and random vectors are denoted as \textit{italic}, boldface, small-case letters \textit{e.g.} $\bm{x}$.
The $n-$th component of a vector $\bm{x}$ is denoted as $x_n$.
A complex vector of length $n$ is defined as $\bm{x} \in \mathbb{C}^{n\times 1}$. All
vectors are column vectors. Matrices are denoted as \textit{italic}, boldface, capital
letters as in $\bm{X}$; the identity matrix of order $n$ is denoted as $\bm{I}_n$.
We define a discrete-time circularly symmetric Gaussian process $\bm{z}$ as any collection
of random varibles $\bm{z}~=~\bm{x}~+~j\bm{y}$, $j \triangleq \sqrt{-1}$, such that
$\bm{x}$ and $\bm{y}$ are iid jointly Gaussian with zero mean vector and covariance
matrix given by $\mathbb{E}\left[\bm{z}\bm{z}^{\dagger}\right]$, in which $\bm{z}^\dagger$
means the conjugate transpose of $\bm{z}$. The expectation value wrt a random variable $x$
is denoted as $\mathbb{E}_x$. The probability of an event $A$ is denoted as $\mathbb{P}(A)$.
The indicator function is denoted as $\mathbb{I}(\cdot)$ and it evaluates to one if its
argument is true and zero otherwise.

@@

\textit{A parte problem\'atica aqui \'e encontrar um limitante inferior
para a fun\c c\~ao Gamma. Talvez serie de Taylor resolva.}

@@

Consider the system
\begin{align}
    \bm{y} = h\bm{s} + \bm{w}
\end{align}
in which $\bm{s} \in \mathbb{C}^{n\times 1}$ is a complex On-Off Keying (OOK) signal,
$h$ is a complex $\alpha-\mu$ random variable and $\bm{w}$ is a complex Gaussian process
with zero mean vector and covariance matrix equals $\sigma^2\bm{I}_n$, and $\bm{y}$ is
the received complex vector signal.

Assume that the OOK symbols are equiprobable and that there exist no interference
between the in-phase and quadrature components, then the probability of one bit error
is given as
\begin{align}
    p_{e} = \dfrac{1}{2}\left(\mathbb{P}\left(\hat{y}_i = 0 | s_{i} = 1\right)
                            + \mathbb{P}\left(\hat{y}_i = 1 | s_{i} = 0\right)\right).
\end{align}

Assume that the decoded vector $\bm{\hat{y}}$ is estimated using the minimum distance decoding
rule, i.e.,
